Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-05-05 Paper_v1.tex
Small changes

\section{Introduction}
Cell-free protein expression has become a widely used research tool in systems and synthetic biology, and a promising technology for personalized protein production.
From the perspective of research, cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Central amongst these, is direct access to metabolites and the biosynthetic machinery without the interference of a cell wall, or complications associated with cell growth.
This allows us to interrogate the chemical environment while the biosynthetic machinery is operating, potentially at a fine time resolution.
Cell-free protein synthesis (CFPS) systems are arguably the most prominent examples of cell-free systems used today \cite{Jewett:2008aa}.
However, CFPS is not new; CFPS in crude \textit{E.~coli} extracts has been used since the 1960s to explore fundamental biological mechanisms \cite{MATTHAEI:1961aa,NIRENBERG:1961aa}.
Today, cell-free systems are used in a variety of applications ranging from therapeutic protein production \cite{Lu:2014aa} to synthetic biology \cite{Hodgman:2012aa}.
However, if CFPS is to become a mainstream technology for advanced applications such as point of care manufacturing \cite{Pardee2016248}, we must first understand the performance limits and cost of these systems \cite{Jewett:2008aa}.
One tool to address these questions is mathematical modeling.

@@

\subsection{Model derivation and validation}
The cell-free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa},
and adding deletions associated with the specific cell-free system (see materials and methods).
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 264 reactions and 146 species, is shown in Fig.~\ref{fig:network}A.
The network described the major carbon and energy pathways, as well as amino acid biosynthesis and degradation pathways.
Using this network, in combination with effective promoter models taken from Moon et al. \cite{Moon:2012ab}, and literature values for cell-free culture parameters (Table ~\ref{tbl:parameters}),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different \textit{E.~coli} cell-free extracts.
The cell-free metabolic network, along with all model code and parameters, can be downloaded under an MIT software license from the Varnerlab website \cite{varnerlab}.

>>>>

\section{Introduction}
Cell-free protein expression has become a widely used research tool in systems and synthetic biology, and a promising technology for personalized protein production.
From the perspective of research, cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Central amongst these, is direct access to metabolites and the biosynthetic machinery without the interference of a cell wall, or complications associated with cell growth.
This allows us to interrogate the chemical environment while the biosynthetic machinery is operating, potentially at a fine time resolution.
Cell-free protein synthesis (CFPS) systems are arguably the most prominent examples of cell-free systems used today \cite{Jewett:2008aa}.
However, CFPS is not new; CFPS in crude \textit{E.~coli} extracts has been used since the 1960s to explore fundamental biological mechanisms \cite{MATTHAEI:1961aa,NIRENBERG:1961aa}.
Today, cell-free systems are used in a variety of applications ranging from therapeutic protein production \cite{Lu:2014aa} to synthetic biology \cite{Hodgman:2012aa}.
However, if CFPS is to become a mainstream technology for advanced applications such as point of care manufacturing \cite{Pardee2016248}, we must first understand the performance limits and cost of these systems \cite{Jewett:2008aa}.
One tool to address these questions is constraint based modeling.

@@

\subsection{Model derivation and validation}
The cell-free stoichiometric network was constructed by removing growth associated reactions from the \textit{i}AF1260 reconstruction of K-12 MG1655 \textit{E.~coli} \cite{Feist:2007aa},
and adding deletions associated with the specific cell-free system (see materials and methods).
We then added the transcription and translation template reactions of Allen and Palsson for the specific proteins of interest \cite{Allen:2003aa}.
A schematic of the metabolic network, which consisted of 264 reactions and 146 species, is shown in Fig.~\ref{fig:network}A.
The network described the major carbon and energy pathways, as well as amino acid biosynthesis and degradation pathways.
Using this network, in combination with effective promoter models taken from Moon et al. \cite{Moon:2012ab}, and literature values for cell-free culture parameters (Table ~\ref{tbl:parameters}),
we simulated the sequence specific production of two model proteins, chloramphenicol acetyltransferase (CAT) and dual emission green fluorescent protein (deGFP) using
different \textit{E.~coli} cell-free extracts.
In particular, we calculated the transcription rate using effective promoter models in combination with a maximum transcription rate estimated from literature.
We then maximized the rate of translation, which was bounded from above by an estimate of the maximum translation rate from literature.
The parameters in the transcription and translation models were largely derived from literature (Table ~\ref{tbl:parameters}).
Both the transcription and translation rates were subject to resource constraints encoded by the metabolic network.
The cell-free metabolic network, along with all model code and parameters, can be downloaded under an MIT software license from the Varnerlab website \cite{varnerlab}.
